---
title: My dynamic report
output: html_fragment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)
suppressPackageStartupMessages(library(tidyverse))
```

```{css}
head {
  <link href="https://fonts.googleapis.com" rel="stylesheet">

}

body {
        background-color: white;
        font-family: 'Indie Flower', sans-serif;
    }
    h2 {
        color: navy;
        margin-left: 20px;
        font-family: "Times New Roman", serif;
    }
    h3 {
        color: skyblue;
        margin-left: 20px;
        font-family: "Times New Roman", serif;
}
```

## Congestion status of NX2 buses from Smales Farm to University of Auckland

This data records congestion status on NX2 buses from Smales Farm to UoA throughout the week. Looking into the dataframe's structure we see 20 bus trips and 5 columns: time of the record, full seats status, time period of the bus, number of people standing, and the weekday of the bus trip. We have 20 bus trips recorded and 5 columns. 

```{r}
logged_data <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vTM_q8Spw55ImrUv6zcjuWx5TqmZ-B1x5kAsCCM9eK5RK1P9tYjlTIO_KIqYeMajrH7McnLzPHBtJI6/pub?gid=378957877&single=true&output=csv") 

latest_data <- rename(logged_data,
        seat_full = 2,
         time = 3,
         number_standing = 4,
         weekday = 5
    
  )

glimpse(latest_data)
echo=FALSE
```
### Summary statistics
We only have one numeric variable and that is from the number of people standing. Let's look at the summary statistics for this variable:
```{r}
average_stand <- latest_data$number_standing %>%
  mean(na.rm = TRUE) %>%
  round(1)
median_stand <- latest_data$number_standing %>%
  median(na.rm = TRUE)
max_stand <- latest_data$number_standing %>%
  max(na.rm = TRUE)
min_stand <- latest_data$number_standing %>%
  min(na.rm = TRUE)

paste("===SUMMARY STATISTICS ON NUMBER OF PEOPLE THAT STANDS===")
paste("The average number of people that stand are: ", average_stand)
paste("The minimum number of people that stand are: ", min_stand)
paste("The maximum number of people that stands are: ", max_stand)

echo=FALSE
```
On average, there are around more than 8 people that have to stand, so the bus tends to carry 8 more people than the seats number. There are bus trips where everyone has a seat but there are times when 25 people must stand, 25 is consider quite a large number of people.
The main insight here is that, in the sample, NX2 buses tends to carry more people than it planned, however a surplus of 8 people can fit on a bus and not worth investing another empty bus to accomodate them. 

### Discover distributions
The bar chart belows shows us the distributions of the other 3 variables, besides number of people standing. This will give us a whole complete picture of what the data shows. 

```{r fig.width = 12, fig.height = 4}
latest_data %>%
  ggplot() +
  geom_bar(aes(x = seat_full),
           fill = "#FFFA82") + 
  labs(title = "Congestion status of NX2 buses",
       subtitle = "Data collected over period of more than 1 week",
       x = "Congestion status",
       y = "Number of occurences")
```

The most common 2 statuses are contradictory: there're still seats left and there's a crowd that have to stand. We can see that the number of times that there is a crowd standing is equal to number of times where the bus is not full. 
This shows that large surpluses are abnormally common. Suggesting need to increase bus trips to accomodate surpluses.

```{r fig.width = 12, fig.height = 4}
latest_data %>%
  ggplot() +
  geom_bar(aes(x = weekday),
           fill = "#00FFFF") + 
  labs(title = "Weekdays that I was on the NX2",
       subtitle = "Data collected over period of more than 1 week",
       x = "Weekday",
       y = "Number of occurences")


```

This sample collects data only on Monday to Thursday, with an even share of weekdays. However, the data can not represent Friday's statuses.
```{r}
latest_data %>%
  ggplot() +
  geom_bar(aes(x = time),
           fill = "#FAC898") + 
  labs(title = "Time periods that I was on the NX2",
       subtitle = "Data collected over period of more than 1 week",
       x = "Time period",
       y = "Number of occurences")


```

This sample collects most data in the morning, which explains the large surplus of people, since everyone needs to get to school and to work at the same time, creating very high traffic for buses. 
This reveals the need to investigate further in bus traffics in the morning.
